---
title: "PML_PA"
author: "AS"
date: "Tuesday, October 21, 2014"
output: html_document
---

```{r echo = FALSE}
setwd("C:/Users/Dell/Desktop/Coursera/Practical Machine Learning")
library(caret)
library(randomForest)
library(e1071)


```

# Load  Data Set
```{r  echo = TRUE}
orig_data_train <- read.csv("pml-training.csv");
orig_data_test <-  read.csv("pml-testing.csv");
```

# Review the train data
```{r results ='hold'}

str(orig_data_train)

```


# remove irrelevant variables from both data sets

```{r}
irInd <- grep("X|timestamp|user_name|new_window", names(orig_data_train));
orig_data_train <- orig_data_train[,-irInd];
orig_data_test <- orig_data_test[,-irInd];

```
# We have to deal with NAs. we will assign zero to all NAs

```{r }
orig_data_train[is.na(orig_data_train)] <- 0
orig_data_test[is.na(orig_data_test)] <- 0
```
# Now we will remove near zero variance from both data sets
```{r}
nrInd <- nearZeroVar(orig_data_train);
orig_data_train <- orig_data_train[,-nrInd];
orig_data_test <- orig_data_test[,-nrInd];
```

# Performing partitioning with caret package  
Use 70% of data for training. Save rest of data for test.  
```{r}
set.seed(222);
cdptrain <- createDataPartition(orig_data_train$classe, p=0.7, list = FALSE);
new_data_train <- orig_data_train[cdptrain,];
newvalid <- orig_data_train[-cdptrain,];


library(randomForest)
```


# Fitting the Model with randomForest package to train data
```{r}
modelFit <- randomForest(classe ~., data=new_data_train, do.trace = 10)
modelFit
```

# Fitting the model with randomForest package to test data for predicting

```{r}
modPred <- predict(modelFit,newvalid);
confusionMatrix(modPred, newvalid$classe);
```

#Results

#Now running the model on the test set (20 test cases).  These predictions are submitted as part of the assignment.

```{r}
predFinal <- predict(modelFit, orig_data_test);
predFinal;

```
#The algorithm does correctly predicts the way in which the exercises were carried out.

# These outputs are to be saved to files for submission.
```{r}
results <- as.vector(predFinal)

pml_write_files = function(x) {
  n = length(x)
  for (i in 1:n) {
    filename = paste0("problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote=FALSE, row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(results)

```